---
title: "Hypothesis Testing: Comparing Two Groups"
author: "Seth"
date: "`r Sys.time()`"
output: html_document
---

Consider the Flight Delays dataset discussed in Chapter 1 of the Chihara & Hesterberg textbook, which contains data on selected flights departing from LaGuardia airport in 2009. Complete all problems below using that dataset. 

```{r message = FALSE}
delays <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Chihara/FlightDelays.csv")
str(delays)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(BSDA)
```


### Problem 1

Conduct a traditional t-test and a permutation resampling test to determine whether the average delay times of the two airlines differ. 

```{r message = FALSE}
summaries <- delays %>% group_by(Carrier) %>% summarize(n = n(), xbar = mean(Delay), stdv = sd(Delay))
print.data.frame(summaries)
ggplot(delays, aes(x = Carrier, y = Delay)) + geom_boxplot()
ggplot(delays, aes(sample = Delay, color = Carrier)) + stat_qq() + stat_qq_line()
# Ho: Mean delay time of flight 1 = Mean delay time of flight 2 
obsMeanDiff <- -diff(summaries$xbar)
print(obsMeanDiff)
delayTest<- t.test(delays$Delay ~ delays$Carrier, var.equal = FALSE)
print(delayTest)
```

```{r}
N <- 10^5 - 1
permMeanDiff <- numeric(N)

testdata <- delays$Delay

for (i in 1:N){
   index <- sample(4029, 403)
   AA <- testdata[index]
   UA <- testdata[-index]
   permMeanDiff[i] <- mean(AA) - mean(UA)
}

ggplot(NULL, aes(x = permMeanDiff)) + 
  geom_histogram(bins = 30, fill = "lightblue") + 
  geom_vline(xintercept = obsMeanDiff, color = "red") +
  labs(title = "Mean Differences for 1,000,000 Resamples of the Data")

# lower tail test with Ha: mu_b < mu_g

pvalue.lower <- (sum(permMeanDiff <= obsMeanDiff) + 1) / (N + 1)

# upper tail test with Ha: mu_b > mu_g

pvalue.upper <- (sum(permMeanDiff >= obsMeanDiff) + 1) / (N + 1)

# two-tail test with Ha: mu_b =/= mu_g

pvalue.twotail <- 2 * min(pvalue.lower, pvalue.upper)

data.frame(alternative = c("Lower-Tail", "Upper-Tail", "Two-Tailed"),
           pvalue      = c(pvalue.lower, pvalue.upper, pvalue.twotail))
```

### Problem 2

Conduct a traditional z-test and a permutation resampling test to determine whether United Airlines has a greater proportion of flights delayed by 30 minutes than American Airlines. 

```{r}
summaries <- delays %>% 
  group_by(Carrier) %>% summarize(n = n(), 
                                phat_y = mean(Delayed30 == "No"), 
                                phat_n = mean(Delayed30 == "Yes"))
print.data.frame(summaries, rwo.names = FALSE)
table(delays$Carrier, delays$Delayed30)

obsPropDiff_n <- -diff(summaries$phat_n)
obsPropDiff_y <- -diff(summaries$phat_y)
print(obsPropDiff_n)
print(obsPropDiff_y)
homeTest1 <- prop.test(table(delays$Carrier, delays$Delayed30), correct = FALSE)
print(homeTest1)

homeTest2 <- prop.test(table(delays$Carrier, delays$Delayed30), correct = TRUE)
print(homeTest2)
```

```{r}
N <- 10^5 - 1
permPropDiff <- numeric(N)

testdata <- delays$Delayed30

for (i in 1:N){
   index <- sample(4029, 403)
   AA <- testdata[index]
   UA  <- testdata[-index]
   permPropDiff[i] <- mean(UA == "yes") - mean(AA == "yes")
}

ggplot(NULL, aes(x = permPropDiff)) + 
  geom_histogram(bins = 30, fill = "lightblue") + 
  geom_vline(xintercept = obsPropDiff_y, color = "red") +
  labs(title = "Resampling Mean Differences")

# lower tail test with Ha: mu_b < mu_g

pvalue.lower <- (sum(permPropDiff <= obsPropDiff_y) + 1) / (N + 1)

# upper tail test with Ha: mu_b > mu_g

pvalue.upper <- (sum(permPropDiff >= obsPropDiff_y) + 1) / (N + 1)

# two-tail test with Ha: mu_b =/= mu_g

pvalue.twotail <- 2 * min(pvalue.lower, pvalue.upper)

data.frame(alternative = c("Lower-Tail", "Upper-Tail", "Two-Tailed"),
           pvalue      = c(pvalue.lower, pvalue.upper, pvalue.twotail))
```

### Problem 3

Conduct a traditional F-test and permutation resampling test to determine whether the variance in flight delay times is different between the two carriers. Unlike means and proportions, the test statistic used for a test of two variances is the _ratio_ of the variances, not the difference.

```{r}
summariesF <- delays %>% group_by(Carrier) %>% summarize(n = n(), varience = var(DepartTime), stdv = sd(DepartTime))
obsVarDiff <- -diff(summariesF$varience)
var.test(delays$DepartTime ~ delays$Carrier,
         alternative = "two.sided")
print(summariesF)
```

```{r}
# N <- 10^5 - 1
# permVarDiff <- numeric(N)

# testdata <- delays$DepartTime

# for (i in 1:N){
#   index <- sample(4029, 403)
#   AA <- testdata[index]
#   UA <- testdata[-index]
#   permVarDiff[i] <- var(AA)/var(UA)
#}

# ggplot(NULL, aes(x = permVarDiff)) + 
#  geom_histogram(bins = 30, fill = "lightblue") + 
#  geom_vline(xintercept = obsVarDiff, color = "red") +
#  labs(title = "Mean Differences for 1,000,000 Resamples of the Data")

# lower tail test with Ha: mu_b < mu_g

# pvalue.lower <- (sum(permVarDiff <= obsVarDiff) + 1) / (N + 1)

# upper tail test with Ha: mu_b > mu_g

# pvalue.upper <- (sum(permVarDiff >= obsVarDiff) + 1) / (N + 1)

# two-tail test with Ha: mu_b =/= mu_g

# pvalue.twotail <- 2 * min(pvalue.lower, pvalue.upper)

# data.frame(alternative = c("Lower-Tail", "Upper-Tail", "Two-Tailed"),
#           pvalue      = c(pvalue.lower, pvalue.upper, pvalue.twotail))
```
<hr>

END!!!